---
title: 'Gates & Mydlarz et al: 2014 - 2016 bleaching and recovery of Montipora capitata'
author: "Chris Wall"
date: "7/7/2017"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '4'
---
```{r setup, message=FALSE, warning=FALSE, ECHO=FALSE}
## load packages
library(lmerTest)
library(lmtest)
library(lme4)
library(car)
library(effects)
library(gplots)
library(plotrix)
library(psych)
library(ggplot2)
library(grid)
library(gridExtra)
library(scales)
library(MASS)
library(lsmeans)
library(pbkrtest)
library(cowplot)
library(devtools)
library(dplyr)
#install_github("vqv/ggbiplot")
#library(ggbiplot)

```
## Background
Data is from Kāne'ohe Bay, O'ahu, Hawai'i. Data extends across **two archipelagic bleaching events and the subsequent recovery periods** where tempertaure stress subsided. Data periods represent:  
- first bleaching (October 2014) 
- first recovery (February 2015) 
- second bleaching (October 2015)  
- second recovery (February 2016)  

Physiological and immunological parameters were collected from from *Montipora capitata* collected from two locations (Sites):  
- Lilipuna -- a shallow fringing reef west of the Hawaii Insititute of Marine Biology  
- Reef 14 -- a shallow inshore patch reef in central Kāne'ohe Bay

**Lilipuna** has been categorized as experiencing near shore impacts from freshwater inundation, subteranean groundwater discharge, and seawater biogeochemistry influeces by long residence times, low flow, and riverine/terrigenous nutrient inputs. Additionally, pCO2 flux at this site is "moderate", with on average ~200ppm pCO2 +/- flux

**Reef 14** has seawater with a shorter residence time, little terrestrial impact, but has a greater pCO2 flux due to reef metabolism. This flux can be substantial, being > 600 ppm pCO2 in a 24h period and reaching maximum values of >900ppm pCO2.

## Import and Normalize
Raw data from physiological assessments is imported and normalized according to established protocols. Immunolgical data has been normalized previously and is represented as the final dataframe.

```{r}
###########################################################################
## data period: October 2014, February 2015, October 2015, February 2016 ##
###########################################################################

# working directory
getwd()
rm(list=ls())

### data file
# all lab data from all time points (physio and immuno, PLUS color analysis)
physimmun<-read.csv("data/Gates_Mydlarz_20142016_physimmun.csv", header=T, na.string=NA) #physimmuno data
qPCR.data<-read.csv("data/Gates_Mydlarz_20142016_qPCR.csv", header=T, na.string=NA) #qPCR data

data<-read.csv("data/Gates_Mydlarz_20142016_ALL_DATA.csv", header=T, na.string=NA) #qPCR data


################################################
# PHYSIOLOGY: calculate, normalize dependent variables
################################################
data$cells..ml<-as.numeric(data$cells..ml)
data$ID<-as.factor(data$ID)


# helpful shorthand
SA<-data$surface.area.cm2 # surface area in cm2
blastate<-data$total.blastate.ml # tissue slurry blastate in ml

# Symbiodinium.cells..cm2 == cell.ml * blastate / SA
data$zoox<- (data$cells..ml*blastate)/SA

# ug.chlorophyll.a..cm2 == ug.chl.a.ml * blastate / SA
data$chla<- (data$ug.chla..ml*blastate)/SA

# pg.chlorophyll.a..cell == ug.chl.a.ml * 10^6 / cells.ml
data$chlacell<- (data$ug.chla..ml*10^6)/data$cells..ml

# AFDW.mg..cm2 == convert AFDW g to mg, mutiply by blastate volume, divide by cm2
data$AFDW<- (data$g.AFDW..ml*1000*blastate)/SA

# mg.protein..cm2 == mg.protein.ml * blastate / SA
data$prot<- (data$mg.prot..ml*blastate)/SA


#### rearrange and clean up

# drop unwanted columns by name
data<-data[!colnames(data) %in% c("total.blastate.ml", "surface.area.cm2", "mg.prot..ml", "cells..ml", "ug.chla..ml", "g.AFDW..ml")]

data<-data[!(data$ID=="N37"), ] # this coral has an NA for symbiont type, remove from df

# reorder columns to show: hierarchy of structure, physio., immuno., colorimetry
# this is now the working dataframe
df<-data[, c(1:6, 20:24, 7:19)]

# remove negative values for MEL and POX and
df$MEL[df$MEL <= 0]<- NA
df$MEL[df$MEL > 0.05]<- NA

df$POX[df$POX <= 0]<- NA
df$CAT[df$CAT <= 0]<- NA
df$PPO[df$PPO <= 0]<- NA


df$chlacell[df$chlacell >= 10]<- NA # removing 2 outliers
df$CAT[df$CAT >= 900]<- NA # 2 outliers removed


# don't need "Ramp" lab data, this was lab thermally stressed. Remove this.
df<-df[!(df$Status=="Lab-Ramp"),]

#write.csv(df, "df.normalized.csv")
```

## Color scores

Overall, color is a poor measure of bleaching in these fragments, and there is not a strong relationship between color (R/G/B) and physiological bleaching quantification. Issues may be in the approach to get color scores (in shooting images, or downstream in photoshop), as well as in the fact the same color score can represent significant changes in cell density or chla content thereby making the relationship noisy (although significant).  
- red and green are best explantory variable for chla (~25 % R2)  
- blue and green slightly better to explain zoox density (~33 % R2)  

```{r, fig.width= 2, fig.height= 2, eval=FALSE}
par(mfrow=c(1,3), mar=c(5,4,1,2), pty="sq")

##########################################
# testing color scores and chla relationship
mod<-lm(coral.red.adj~chla, data=df)
plot(coral.red.adj~chla, data=df)
abline(mod, col="red")

mod<-lm(coral.blue.adj~chla, data=df)
plot(coral.blue.adj~chla, data=df)
abline(mod, col="blue")

mod<-lm(coral.green.adj~chla, data=df)
plot(coral.green.adj~chla, data=df)
abline(mod, col="green")


### Color score by *Symbiodinium* cells cm-2
##########################################
# testing color score and zoox relationship
mod<-lm(coral.red.adj~zoox, data=df)
plot(coral.red.adj~zoox, data=df)
abline(mod, col="red")

mod<-lm(coral.blue.adj~zoox, data=df)
plot(coral.blue.adj~zoox, data=df)
abline(mod, col="blue")

mod<-lm(coral.green.adj~zoox, data=df)
plot(coral.green.adj~zoox, data=df)
abline(mod, col="green")


####### PCA and color score
####### Run the PCA first and save PC1 data 

pca.df<-(df[-c(1:113),c(2:6, 17:19)]) # all data with only "Event" and "Site" as categories
colnames(pca.df)<-c("Period", "Status", "Site", "Status_Site", "ID", "red", "green", "blue")
pca.df<-pca.df[-9, ] #remove NA row

# apply PCA - scale. = TRUE for center as advised
PC <- prcomp(pca.df[, 6:8], center = TRUE, scale.= TRUE)  

#correlatin matrix helps to standardize the data and account for different scales
# print(PC) # to print PC loadings

# summary(PC) #signficance of PCs: proportion of variance captured, cumulative variance
# plot(PC, type="lines") will plot PCs

newdat<-PC$x[,1:3]; # newdata frame with only PCs 
# PC1 explain ~>82% of variance

# save the file of PCs
write.csv(newdat, "PC1-allcolors_alltimes.csv")

# the SDEV^2 is the eignevalue
ev<-PC$sdev^2 # PC1-3 have eignevalues >1
# ev/sum(ev) # to give proportional eignevalues

#########
# plot it as PCA analysis by bleaching period

## PC1 and PC2
g <- ggbiplot(PC, choices = 1:2, obs.scale = 1, var.scale = 1, 
              groups= pca.df[,1], ellipse = TRUE, 
              circle = FALSE) +
              scale_color_discrete(name = '') +
  theme_bw() +
  theme(legend.text=element_text(size=15)) +
  theme(panel.background = element_rect(colour = "black", size=1))+
  theme(legend.key = element_blank())+
  theme(legend.direction = 'horizontal', legend.position = 'top') +theme(aspect.ratio=0.7)
print(g)

##################

### graphing the PC1 color score relationship with chlorophyll or *Symbiodinium* density**
### significant relationships, but only 27% (chla) and 35% (zoox) of variance explained

##########################################
# testing color scores and chla relationship
mod<-lm(PC1~chla, data=df)
plot(PC1~chla, data=df)
abline(mod, col="mediumorchid3")

##########################################
# testing PC1 score and zoox relationship
mod<-lm(PC1~zoox, data=df)
plot(PC1~zoox, data=df)
abline(mod, col="mediumorchid")



### Categorical binning for bleaching
# First, looked at all the data as histograms of chlorophyll a and zoox density across all 4 time periods. Second, looked at the histograms forjust bleaching, and then just recovery periods. While it is difficult to say when a coral is truly bleached, a conservative measure of < 3 ug chla/cm2 or < 1.5 million cells may be used as a relative measure of bleaching.


#############################
# all data: bleaching and recovery periods
par(mfrow=c(1,2), mar=c(5,4,1,2), pty="sq")
hist(df$zoox, main="all time points"); hist(df$chla, main="all time points")


#############################
#### just "bleaching" periods
bleaching.df<-df[df$Status=="Bleaching", ]

par(mfrow=c(1,2), mar=c(5,4,1,2), pty="sq")
hist(bleaching.df$zoox, main="bleaching periods")
hist(bleaching.df$chla, main="bleaching periods") 
# less than 3 ug chla/cm2 is good indication of "bleached"
# less than 1.5 million cells/cm2 is good indication of "bleached

# what about symbiont community
hist(bleaching.df[bleaching.df$dom=="D",]$chla)
hist(bleaching.df[bleaching.df$dom=="C",]$chla)


#############################
#### just "recovery" periods
recovery.df<-df[df$Status=="Recovery", ]

par(mfrow=c(1,2), mar=c(5,4,1,2), pty="sq")
hist(recovery.df$zoox, main="recovery periods")
hist(recovery.df$chla,  main="recovery periods")

### Bleaching categories
# Decided to bin the data based on bleaching (< 40%  ug chla cm-2) and pigmented (> 60% ug chla cm-2). This should help in finding those corals more affected by the thermal stress, and those not affected as severely.

#### applying binning

#### no break here for Lab 2014--these are all effectively field controls for physiology
Field.2014<-df[(df$Period=="2014 Field.Feb"),]
Field.2014<- Field.2014 %>% mutate(
            bleach.category = "Field-Pre")

#### no break here for Lab 2014--these are all effectively field controls for physiology
Lab.2014<-df[(df$Period=="2014 Lab"),]
quantile(Lab.2014$chla, 0.4, na.rm=TRUE) # = 3.86
Lab.2014<- Lab.2014 %>% mutate(
            bleach.category = "Lab-pigmented")

#### 40% quantile for October 2014
Oct.2014<-df[(df$Period=="2014 Oct"),]
quantile(Oct.2014$chla, 0.4, na.rm=TRUE) # = 2.97
Oct.2014<- Oct.2014 %>% mutate(
            bleach.category = ifelse(chla < "2.97", "pale", "pigmented"))

#### 40% quantile for February 2015
Feb.2015<-df[(df$Period=="2015 Feb"),]
quantile(Feb.2015$chla, 0.4, na.rm=TRUE) # = 3.72
Feb.2015<- Feb.2015 %>% mutate(
            bleach.category = ifelse(chla < "3.72", "pale", "pigmented"))

#### 40% quantile for October 2015
Oct.2015<-df[(df$Period=="2015 Oct"),]
quantile(Oct.2015$chla, 0.4, na.rm=TRUE)  # = 2.49
Oct.2015<- Oct.2015 %>% mutate(
            bleach.category = ifelse(chla < "2.49", "pale", "pigmented"))

#### 40% quantile for January 2016
Feb.2016<-df[(df$Period=="2016 Feb"),]
quantile(Feb.2016$chla, 0.4, na.rm=TRUE)  # = 3.84
Feb.2016<- Feb.2016 %>% mutate(
            bleach.category = ifelse(chla < "3.84", "pale", "pigmented"))


##########
## Now combine all the dataframes above back into a single df
df<-rbind(Field.2014, Lab.2014, Oct.2014, Feb.2015, Oct.2015, Feb.2016)

```

### Summary dataframes
These dataframes show mean, SE and n for each group. This can be subset to make figures or exported as a summary file. The header of the dataframe is shown below...

```{r, ECHO=FALSE}
# make summary dataframes

# dataframes of means can later be divided into categories by "C or D" dominated

###------#### means

# physiology
zoox.m<-aggregate(zoox~Site+Status+Period+dom,data=df, mean)
chla.m<-aggregate(chla~Site+Status+Period+dom,data=df, mean)
chlacell.m<-aggregate(chlacell~Site+Status+Period+dom, data=df, mean)
prot.m<-aggregate(prot~Site+Status+Period+dom, data=df, mean)
AFDW.m<-aggregate(AFDW~Site+Status+Period+dom, data=df, mean)

# imunology
CAT.m<-aggregate(CAT~Site+Status+Period+dom,data=df, mean)
POX.m<-aggregate(POX~Site+Status+Period+dom,data=df, mean)
SOD.m<-aggregate(SOD~Site+Status+Period+dom,data=df, mean)
PPO.m<-aggregate(PPO~Site+Status+Period+dom,data=df, mean)
MEL.m<-aggregate(MEL~Site+Status+Period+dom,data=df, mean)

###------#### SE
# physiology
zoox.SE<-aggregate(zoox~Site+Status+Period+dom,data=df, std.error)
chla.SE<-aggregate(chla~Site+Status+Period+dom,data=df, std.error)
chlacell.SE<-aggregate(chlacell~Site+Status+Period+dom, data=df, std.error)
prot.SE<-aggregate(prot~Site+Status+Period+dom, data=df, std.error)
AFDW.SE<-aggregate(AFDW~Site+Status+Period+dom, data=df, std.error)

# imunology
CAT.SE<-aggregate(CAT~Site+Status+Period+dom,data=df, std.error)
POX.SE<-aggregate(POX~Site+Status+Period+dom,data=df, std.error)
SOD.SE<-aggregate(SOD~Site+Status+Period+dom,data=df, std.error)
PPO.SE<-aggregate(PPO~Site+Status+Period+dom,data=df, std.error)
MEL.SE<-aggregate(MEL~Site+Status+Period+dom,data=df, std.error)

###------#### n-value
# physiology
zoox.n<-aggregate(zoox~Site+Status+Period+dom,data=df, length)
chla.n<-aggregate(chla~Site+Status+Period+dom,data=df, length)
chlacell.n<-aggregate(chlacell~Site+Status+Period+dom, data=df, length)
prot.n<-aggregate(prot~Site+Status+Period+dom, data=df, length)
AFDW.n<-aggregate(AFDW~Site+Status+Period+dom, data=df, length)

# imunology
CAT.n<-aggregate(CAT~Site+Status+Period+dom,data=df, length)
POX.n<-aggregate(POX~Site+Status+Period+dom,data=df, length)
SOD.n<-aggregate(SOD~Site+Status+Period+dom,data=df, length)
PPO.n<-aggregate(PPO~Site+Status+Period+dom,data=df, length)
MEL.n<-aggregate(MEL~Site+Status+Period+dom,data=df, length)


#### physio dataframe
fig.df.phys<-cbind(zoox.m, zoox.SE[c(5,0)],chla.m[c(5,0)], chla.SE[c(5,0)], chlacell.m[c(5,0)], chlacell.SE[c(5,0)], prot.m[c(5,0)], prot.SE[c(5,0)], AFDW.m[c(5,0)], AFDW.SE[c(5,0)], chla.n[c(5,0)])

colnames(fig.df.phys)<-c("Site","Status", "Period", "dom", "zoox", "zoox.SE", "chla", "chla.SE", "chlacell", "chlacell.SE", "prot", "prot.SE", "AFDW", "AFDW.SE", "N-chla")

# make an interaction column
fig.df.phys$int<-interaction(fig.df.phys$Site, fig.df.phys$dom)
fig.df.phys$int<-factor(fig.df.phys$int)
####


#### immuno dataframe
fig.df.immuno<-cbind(CAT.m, CAT.SE[c(5,0)], POX.m[c(5,0)], POX.SE[c(5,0)], SOD.m[c(5,0)], SOD.SE[c(5,0)], PPO.m[c(5,0)], PPO.SE[c(5,0)], MEL.m[c(5,0)], MEL.SE[c(5,0)])

colnames(fig.df.immuno)<-c("Site","Status", "Period", "dom", "CAT", "CAT.SE", "POX", "POX.SE", "SOD", "SOD.SE", "PPO", "PPO.SE", "MEL", "MEL.SE")

# make an interaction column
fig.df.immuno$int<-interaction(fig.df.immuno$Site, fig.df.immuno$dom)
fig.df.immuno$int<-factor(fig.df.immuno$int)

# write.csv(xxx, "Figure.df.xxx.csv")
```

## Figures

```{r}
# side by side figures separated by Sites 
### Physiological figures

color.scheme<-c("gray80", "gray60", 'lightskyblue', 'dodgerblue', 
                "gray50", "gray30", "thistle", "coral")
break.points<-c("Lilipuna.Pre.C", "Lilipuna.Pre.D", "Lilipuna.C", "Lilipuna.D", 
                "Reef 14.Pre.C", "Reef 14.Pre.D", "Reef 14.C", "Reef 14.D")
legend.names<-c("Lil.Pre C", "Lil.Pre D", "Lil C", "Lil D", 
                "R14.Pre C", "R14.Pre D", "R14 C", "R14 D")
axis.labels<-c("Feb 2014 \nPre-Bleaching", "Oct 2014 \nBleaching","Feb 2015 \nRecovery","Oct 2015 \nBleaching", "Feb 2016 \nRecovery")


Fig.formatting<-(theme_classic()) +
  theme(text=element_text(size=10),
  axis.line=element_blank(),
  legend.text.align = 0,
  legend.text=element_text(size=10),
    #legend.title = element_blank(),
      panel.border = element_rect(fill=NA, colour = "black", size=1),
      aspect.ratio=1, 
    axis.ticks.length=unit(0.25, "cm"),
    axis.text.y=element_text(
      margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=10), 
    axis.text.x=element_text(
      margin=unit(c(0.5, 0.5, 0.5, 0.5), "cm"), colour="black", size=8)) +
  theme(legend.key.size = unit(0.4, "cm")) +
  theme(aspect.ratio=1) +
  theme(panel.spacing=unit(c(0, 0, 0, 0), "cm"))


pd <- position_dodge(0.15) #offset for error bars


########################
#### graph as category of C/D
######################## 

fig.df.phys$int<-factor(fig.df.phys$int, levels=c("Lilipuna.Pre.C", "Lilipuna.Pre.D", "Lilipuna.C", "Lilipuna.D", "Reef 14.Pre.C", "Reef 14.Pre.D", "Reef 14.C", "Reef 14.D"))
```

**Symbiodinium cm-2**
```{r}
###################
# zoox 
zoox.fig<-ggplot(data=fig.df.phys, aes(x=Period, y=zoox/10^6, group=int, color=int)) + geom_errorbar(aes(ymin=zoox/10^6-zoox.SE/10^6, ymax=zoox/10^6+zoox.SE/10^6), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3, pch=19) +
  coord_cartesian(ylim=c(0, 3)) +
  ylab(expression(paste(~italic("Symbiodinium") ~(10^6~cells~cm^-2), sep=""))) +
  scale_color_manual(values=c(color.scheme),
                     name= "Site:Clade",
                     breaks=break.points,
                     labels=legend.names) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting

zoox.fig

```

**Chlorophyll *a* cm-2**
```{r}
########## chla cm2 ############
chla.fig<-ggplot(data=fig.df.phys, aes(x=Period, y=chla, group=int, color=int)) + geom_errorbar(aes(ymin=chla-chla.SE, ymax=chla+chla.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 8)) +
  ylab(expression(paste("Chlorophyll", ~italic("a"), ~(mu*g~cm^-2), sep=""))) +
  scale_color_manual(values=c(color.scheme),
                     name= "Site:Clade",
                     breaks=break.points,
                     labels=legend.names) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
chla.fig
```

**Chlorophyll *a* cell-1**
```{r, message=FALSE}
########## chla cell ############
chlacell.fig<-ggplot(data=fig.df.phys, aes(x=Period, y=chlacell, group=int, color=int)) + geom_errorbar(aes(ymin=chlacell-chlacell.SE, ymax=chlacell+chlacell.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 8)) +
  ylab(expression(paste("Chlorophyll", ~italic("a"), ~(pg~cell^-1), sep=""))) +
  scale_color_manual(values=c(color.scheme),
                     name= "Site:Clade",
                     breaks=break.points,
                     labels=legend.names) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
chlacell.fig
```

**Protein biomass cm-2**
```{r}

########## protein ############
prot.fig<-ggplot(data=fig.df.phys, aes(x=Period, y=prot, group=int, color=int)) + geom_errorbar(aes(ymin=prot-prot.SE, ymax=prot+prot.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 1)) +
  ylab(expression(paste("Protein", ~(mg~cm^-2), sep=""))) +
  scale_color_manual(values=c(color.scheme),
                     name= "Site:Clade",
                     breaks=break.points,
                     labels=legend.names) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
prot.fig
```

**Total biomass (AFDW) cm-2**
```{r}

########## AFDW ############
AFDW.fig<-ggplot(data=fig.df.phys, aes(x=Period, y=AFDW, group=int, color=int)) + geom_errorbar(aes(ymin=AFDW-AFDW.SE, ymax=AFDW+AFDW.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 60)) +
  ylab(expression(paste("Total biomass", ~(mg~cm^-2), sep=""))) +
  scale_color_manual(values=c(color.scheme),
                     name= "Site:Clade",
                     breaks=break.points,
                     labels=legend.names) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
AFDW.fig
```

### Imunological figures

**Prophenoloxidase (PPO)**
```{r}
#names(fig.df.immuno)

# site means
CAT.m2<-aggregate(CAT~Site+Status+Period,data=df, mean)
POX.m2<-aggregate(POX~Site+Status+Period,data=df, mean)
SOD.m2<-aggregate(SOD~Site+Status+Period,data=df, mean)
PPO.m2<-aggregate(PPO~Site+Status+Period,data=df, mean)
MEL.m2<-aggregate(MEL~Site+Status+Period,data=df, mean)

###------#### SE
# imunology
CAT.SE2<-aggregate(CAT~Site+Status+Period,data=df, std.error)
POX.SE2<-aggregate(POX~Site+Status+Period,data=df, std.error)
SOD.SE2<-aggregate(SOD~Site+Status+Period,data=df, std.error)
PPO.SE2<-aggregate(PPO~Site+Status+Period,data=df, std.error)
MEL.SE2<-aggregate(MEL~Site+Status+Period,data=df, std.error)

immuno.site.df<-cbind(CAT.m2, CAT.SE2[c(4,0)], POX.m2[c(4,0)], POX.SE2[c(4,0)], SOD.m2[c(4,0)], SOD.SE2[c(4,0)], PPO.m2[c(4,0)], PPO.SE2[c(4,0)], MEL.m2[c(4,0)], MEL.SE2[c(4,0)])

colnames(immuno.site.df)<-c("Site","Status", "Period", "CAT", "CAT.SE", "POX", "POX.SE", "SOD", "SOD.SE", "PPO", "PPO.SE", "MEL", "MEL.SE")

Pre.immuno<-immuno.site.df[c(1:2),]
Pre.immuno$dom<-"unident"
Pre.immuno<-Pre.immuno[, c(1:3, 14, 4:13)]
Pre.immuno$int<-interaction(Pre.immuno$Site, Pre.immuno$dom)

## now this subset of site means can be overlayed with datafame from Oct 2014- Feb 2016
fig.df.immuno.comp<-rbind(Pre.immuno, fig.df.immuno)


color.scheme2<-c("gray70", 'lightskyblue', 'dodgerblue', 
                "gray40", "thistle", "coral")
break.points.immuno<-c("Lilipuna.Pre.unident", "Lilipuna.C", "Lilipuna.D", 
                "Reef 14.Pre.unident", "Reef 14.C", "Reef 14.D")
legend.names.immuno<-c("Lil.Pre unident.", "Lil C", "Lil D", 
                "R14.Pre unident.", "R14 C", "R14 D")


fig.df.immuno.comp$int<-factor(fig.df.immuno.comp$int, levels=c("Lilipuna.Pre.unident","Lilipuna.C", "Lilipuna.D", "Reef 14.Pre.unident", "Reef 14.C", "Reef 14.D"))


########## Prophenoloxidase (PPO) ############
PPO.fig<-ggplot(data=fig.df.immuno.comp, aes(x=Period, y=PPO, group=int, color=int)) + geom_errorbar(aes(ymin=PPO-PPO.SE, ymax=PPO+PPO.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 3)) +
  ylab(expression(paste("PO Activity", ~(Delta~Abs~"490nm"~min^-1~mg~prot^-1), sep=""))) +
  scale_color_manual(values=c(color.scheme2),
                     name= "Site:Clade",
                     breaks=break.points.immuno,
                     labels=legend.names.immuno) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
PPO.fig
```

**Melanin (MEL)**
```{r}
MEL.lab<-aggregate(MEL~Site+Status+Period,data=data, mean)
MEL.labSE<-aggregate(MEL~Site+Status+Period,data=data, std.error)
MEL.df<-cbind(MEL.lab, MEL.labSE[c(4,0)]); colnames(MEL.df[,4:5])<-c("MEL", "MEL.labSE")
MEL.df$Period<-factor(MEL.df$Period, levels=c("2014 Field.Feb", "2014 Lab", "2014 Oct", "2015 Feb", "2015 Oct", "2016 Feb"))

plot(MEL~Period, data=MEL.df, cex.axis=0.8)
#dev.copy(pdf,"Figures/MEL.all.time.pdf", height=5, width=7, encod="MacRoman")
#dev.off

######## Melanin (MEL) #############
MEL.fig<-ggplot(data=fig.df.immuno.comp, aes(x=Period, y=MEL, group=int, color=int)) + geom_errorbar(aes(ymin=MEL-MEL.SE, ymax=MEL+MEL.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 0.02)) +
  ylab(expression(paste("MEL", ~(Abs~"490nm"~mg~tissue^-1), sep="")))+
  scale_color_manual(values=c(color.scheme2),
                     name= "Site:Clade",
                     breaks=break.points.immuno,
                     labels=legend.names.immuno) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
MEL.fig

ggsave("Figures/MEL.field.pdf", height=5, width=7, encod="MacRoman")
```

**Peroxidase (POX)**
```{r}

########## Peroxidase (POX) ##############
POX.fig<-ggplot(data=fig.df.immuno.comp, aes(x=Period, y=POX, group=int, color=int)) + geom_errorbar(aes(ymin=POX-POX.SE, ymax=POX+POX.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 1.5)) +
  ylab(expression(paste("POX activity", ~(Delta~Abs~"470nm"~min^-1~mg~prot^-1), sep=""))) +
  scale_color_manual(values=c(color.scheme2),
                     name= "Site:Clade",
                     breaks=break.points.immuno,
                     labels=legend.names.immuno) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
POX.fig
```

**Catalase (CAT)**
```{r}

###### Catalase (CAT) ########
CAT.fig<-ggplot(data=fig.df.immuno.comp, aes(x=Period, y=CAT, group=int, color=int)) + geom_errorbar(aes(ymin=CAT-CAT.SE, ymax=CAT+CAT.SE), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 800)) +
  ylab(expression(paste("CAT activity", ~(Abs~"490nm"~mg~tissue^-1), sep="")))+
  scale_color_manual(values=c(color.scheme2),
                     name= "Site:Clade",
                     breaks=break.points.immuno,
                     labels=legend.names.immuno) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
CAT.fig
```

**Superoxide dismutase (SOD)**
```{r}


######### Superoxide dismutase (SOD) ########
SOD.fig<-ggplot(data=fig.df.immuno.comp, aes(x=Period, y=SOD/10^3, group=int, color=int)) + geom_errorbar(aes(ymin=SOD/10^3-SOD.SE/10^3, ymax=SOD/10^3+SOD.SE/10^3), size=.8, width=0, position=pd) +
  geom_line(position=pd, size=.8) +
  geom_point(position=pd, size=3) +
  coord_cartesian(ylim=c(0, 50)) +
  ylab(expression(paste("SOD activity", ~x10^3~mg~protein^-1, sep=""))) +
  scale_color_manual(values=c(color.scheme2),
                     name= "Site:Clade",
                     breaks=break.points.immuno,
                     labels=legend.names.immuno) +
  scale_x_discrete(name ="Year and Event", 
      labels=axis.labels)+
  Fig.formatting
SOD.fig
```


###Models  

Running models of response variables. First check for assumptions of ANOVA and apply transformations where necessary.  
  
*lm(Y~Period x Site x dom, data=model.data, na.action=na.exclude)*  
*Period* is the 4 events, *Site* is the two reefs, *dom* is the symbiont community (C- or D-dominated)

```{r, include=FALSE}
#df ## this is the data

model.data<-df[ , !(names(df) %in% c("coral.red.adj","coral.green.adj", "coral.blue.adj", "PC1", "propC", "propD", "syms"))] # remove unwanted columns

model.data<-model.data[!c(model.data$Period=="2014 Lab"),] 
model.data<-model.data[!c(model.data$Period=="2014 Field.Feb"),] # drop data prior to Oct 2014 bleaching

model.data$Period<-factor(model.data$Period) #re-call as factor to remove dropped levels
model.data$Status<-factor(model.data$Status)


for(i in c(7:9, 11:16)){
  Y<-model.data[,i]
  full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
  R <- resid(full) #save glm residuals
  
  op<-par(mfrow = c(2,2), mar=c(5,4,1,2), pty="sq")
  plot(full, add.smooth = FALSE, which=1)
  QQ <- qqnorm(R, main = colnames(model.data)[i]) 
  QQline <- qqline(R)
  hist(R, xlab="Residuals", main = colnames(model.data)[i])
}

# need BoxCox transformation for most immuno data
Yx<-(model.data$AFDW) 

### test with transformations
boxcox(Yx~Period*Site*dom, data=model.data, lambda=seq(-0.25, 1, length=10))
### test residuals, histogram, QQ
full<-lm(Yx~Period*Site*dom, data=model.data, na.action=na.exclude)
  R <- resid(full) #save glm residuals
  op<-par(mfrow = c(2,2), mar=c(5,4,1,2), pty="sq")
  plot(full, add.smooth = FALSE, which=1)
  QQ <- qqnorm(R) 
  QQline <- qqline(R)
  hist(R, xlab="Residuals")

##### revised model.data (below are data needing transformations)
model.data$CAT<-(model.data$CAT)^0.5
model.data$POX<-(model.data$POX)^0.25
model.data$PPO<-(model.data$PPO) ^0.25
model.data$MEL<-(model.data$MEL)^0.25

## dataframe 'model.data' now ready for full model analysis
```

Run linear models for response variables  

**Symbiodinium cm-2**
```{r symbiont model, warning=FALSE, message=FALSE}
# symbionts ---- 
Y<-model.data$zoox
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site|Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom|Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site*dom|Period)
cld(posthoc, Letters=letters)

plot(allEffects(full), ylab="zoox", par.strip.text=list(cex=0.7))
```

**Chlorophyll *a* cm^-2**
```{r chla model, warning=FALSE, message=FALSE}
# chla ---- 
Y<-model.data$chla
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom|Period)
cld(posthoc, Letters=letters)
plot(allEffects(full), ylab="chla", par.strip.text=list(cex=0.7))
```

**chlorophyll *a* cell-1**
```{r chlcell model, warning=FALSE, message=FALSE}
# chlacell ---- 
Y<-model.data$chlacell
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
##summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom|Period)
cld(posthoc, Letters=letters)
plot(allEffects(full), ylab="AFDW", par.strip.text=list(cex=0.7))
```

**Protein cm-2**
```{r protein model, warning=FALSE, message=FALSE}
# prot ---- 
Y<-model.data$prot
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~dom)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site|Period)
cld(posthoc, Letters=letters)
plot(allEffects(full), ylab="prot", par.strip.text=list(cex=0.7))
```

**Catalase (CAT)**
```{r CAT Model, warning=FALSE, message=FALSE}
######################## 
### immunology
########################

# CAT ---- 
Y<-model.data$CAT
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site|Period)
cld(posthoc, Letters=letters)
plot(allEffects(full), ylab="CAT", par.strip.text=list(cex=0.7))
```

**Peroxidase (POX)**
```{r POX model, warning=FALSE, message=FALSE}
# POX ---- 
Y<-model.data$POX
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom|Period)
cld(posthoc, Letters=letters)

plot(allEffects(full), ylab="POX", par.strip.text=list(cex=0.7))
```

**Superoxide dismutase (SOD)**
```{r SOD model, warning=FALSE, message=FALSE}
# SOD ---- 
Y<-model.data$SOD
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site)
cld(posthoc, Letters=letters)

plot(allEffects(full), ylab="SOD", par.strip.text=list(cex=0.7))
```

**Prophenoloxidase (PPO)**
```{r PPO model, warning=FALSE, message=FALSE}
# PPO ---- 
Y<-model.data$PPO
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~dom)
cld(posthoc, Letters=letters)
plot(allEffects(full), ylab="PPO", par.strip.text=list(cex=0.7))
```

**Melanin (MEL)**
```{r MEL model, warning=FALSE, message=FALSE}
# MEL ---- 
Y<-model.data$MEL
full<-lm(Y~Period*Site*dom, data=model.data, na.action=na.exclude)
#summary(full)
print(Anova(full, type=2), digits=4)

posthoc<-lsmeans(full, ~Period)
cld(posthoc, Letters=letters)

posthoc<-lsmeans(full, ~Site|Period)
cld(posthoc, Letters=letters)
plot(allEffects(full), ylab="prot", par.strip.text=list(cex=0.7))
```

```{r, MultiVariate}
# multivariate test using adonis-- Manova, mutivariate analysis of variance
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
require(graphics)
library(tidyr)
library(vegan)
library(RVAideMemoire)

df.MV<-df

# remove columns unnecessary for final analysis, few factors retained
df.MV<-df.MV[ , !names(df.MV) %in% c("Species", "Status_Site", "ID", "coral.red.adj", "coral.green.adj", "coral.blue.adj", "PC1", "propC", "propD", "syms")]

# remove prior-to-bleaching timepoints and unused levels that arise from this action
df.MV<-df.MV[(!df.MV$Period=="2014 Lab"),] 
df.MV<-df.MV[(!df.MV$Period=="2014 Field.Feb"),]
df.MV$Period<-droplevels(df.MV$Period) 
df.MV$Status<-droplevels(df.MV$Status)

#drop all NAs, removing rows where NA was observed
df.MV.noNA<-df.MV %>% drop_na()

# need to make 2 dataframes, one with only data (no factors) and another with factors+data

######## Lilipuna only #######
Manova.Lil.lev.dat<-df.MV.noNA %>% # "lev.dat" has levels and data
  filter(Site=="Lilipuna") 

Manova.Lil.dat<-Manova.Lil.lev.dat %>% # "dat" is just data
  select(-c(Period, Status, Site, dom)) #  sans factors

##  Run MANOVA Lilipuna
MVA.Lilipuna<-adonis2(Manova.Lil.dat~Period*dom, data=Manova.Lil.lev.dat,permutations=1000, 
                method="bray", sqrt.dist = TRUE)
MVA.Lilipuna

NMDS.Lil<-metaMDS(Manova.Lil.dat, distane='bray', k=2, trymax=100) 
stressplot(NMDS.Lil)

factor.df.Lil<-Manova.Lil.lev.dat %>% 
  select(c(Period, Site, Status, dom)) # get just the factors

NMDS.df.Lil<-data.frame(x=NMDS.Lil$point[,1], y=NMDS.Lil$point[,2], 
                    Period=as.factor(factor.df.Lil[,1]),
                    Site=as.factor(factor.df.Lil[,2]),
                    Status=as.factor(factor.df.Lil[,3]),
                    dom=as.factor(factor.df.Lil[,4]))

colnames(NMDS.df.Lil)[1:2]<-c("MDS1", "MDS2")

# vectors for variables
vars<-envfit(NMDS.Lil, Manova.Lil.dat, permu=999)
#scores(vars, "vectors")

###### plot it
# example of quick and dirty plot in ggplot
ggplot(NMDS.df.Lil, aes(x=MDS1, y=MDS2, col=Period:dom)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

ggplot(NMDS.df.Lil, aes(x=MDS1, y=MDS2, col=Status)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

ggplot(NMDS.df.Lil, aes(x=MDS1, y=MDS2, col=dom)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

ggplot(NMDS.df.Lil, aes(x=MDS1, y=MDS2, col=Status:dom)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

par(mfcol=c(1,2), mar=c(4,4,1,1))


###### plot it
MDS.levels<-levels(NMDS.df.Lil$Period:NMDS.df.Lil$dom)
MDS.lev.leg<-c("C-Bleach '14", "D-Bleach '14", "C-Recov '15", "D-Recov '15", "C-Bleach '15", "D-Bleach '15", "C-Recov '16", "D-Recov '16")
MDS.colors<-c("lightpink", "firebrick2", "slategray1", "dodgerblue2","khaki", "darkorange", "mediumseagreen", "olivedrab")

NMDS.plot<-ordiplot(NMDS.Lil, type="n", main=substitute(paste(italic("Lilipuna"))), cex.main=1, display="sites", cex.lab=0.8, cex.axis=0.8, ylim=c(-0.4, 0.4), xlim=c(-0.6, 0.5))
axis(side = 1, labels = FALSE, tck = -0.03)
abline(h = 0, lty = "dotted")
abline(v = 0, lty = "dotted")
points(NMDS.plot, "sites", cex=0.8, pch=16, col=MDS.colors)
ordiellipse(NMDS.plot, groups=NMDS.df.Lil$Period:NMDS.df.Lil$dom, draw="polygon", col=MDS.colors, alpha=150, kind="sd", conf=0.8)
legend("topright", legend=MDS.lev.leg, inset=c(0.16, -0.02), x.intersp=0.8, cex=0.6, pch=16, col=MDS.colors, pt.cex=0.7, bty="n")
par.new=T
plot(vars, p.max=0.05, cex=0.8, lwd=1)

dev.copy(pdf, "figures/NMDS_Lil.pdf", height=4, width=5, useDingbats=FALSE)
dev.off() 





######## Reef 14 only #######
Manova.R14.lev.dat<-df.MV.noNA %>% # "lev.dat" has levels and data
  filter(Site=="Reef 14") 

Manova.R14.dat<-Manova.R14.lev.dat %>% # "dat" is just data
  select(-c(Period, Status, Site, dom)) #  sans factors

##  Run MANOVA Reef 14
MVA.Reef14<-adonis2(Manova.R14.dat~Period*dom, data=Manova.R14.lev.dat,permutations=1000, 
                method="bray", sqrt.dist = TRUE)
MVA.Reef14

NMDS.R14<-metaMDS(Manova.R14.dat, distane='bray', k=2, trymax=100) 
stressplot(NMDS.R14)

factor.df.R14<-Manova.R14.lev.dat %>% 
  select(c(Period, Site, Status, dom)) # get just the factors

NMDS.df.R14<-data.frame(x=NMDS.R14$point[,1], y=NMDS.R14$point[,2], 
                    Period=as.factor(factor.df.R14[,1]),
                    Site=as.factor(factor.df.R14[,2]),
                    Status=as.factor(factor.df.R14[,3]),
                    dom=as.factor(factor.df.R14[,4]))

colnames(NMDS.df.R14)[1:2]<-c("MDS1", "MDS2")

# vectors for variables
vars<-envfit(NMDS.R14, Manova.R14.dat, permu=999)
#scores(vars, "vectors")

###### plot it
# example of quick and dirty plot in ggplot
ggplot(NMDS.df.R14, aes(x=MDS1, y=MDS2, col=Period:dom)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

ggplot(NMDS.df.R14, aes(x=MDS1, y=MDS2, col=Status)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

ggplot(NMDS.df.R14, aes(x=MDS1, y=MDS2, col=dom)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")

ggplot(NMDS.df.R14, aes(x=MDS1, y=MDS2, col=Status:dom)) +
    geom_point() +
    stat_ellipse() +
    theme_bw() +
    labs(title = "NMDS Plot")


###### plot it
MDS.levels<-levels(NMDS.df.R14$Period:NMDS.df.R14$dom)
MDS.lev.leg<-c("C-Bleach '14", "D-Bleach '14", "C-Recov '15", "D-Recov '15", "C-Bleach '15", "D-Bleach '15", "C-Recov '16", "D-Recov '16")
MDS.colors<-c("lightpink", "firebrick2", "slategray1", "dodgerblue2","khaki", "darkorange", "mediumseagreen", "olivedrab")

NMDS.plot<-ordiplot(NMDS.R14, type="n", main=substitute(paste(italic("Reef 14"))), cex.main=1, display="sites", cex.lab=0.8, cex.axis=0.8, ylim=c(-0.4, 0.4), xlim=c(-0.6, 0.5))
axis(side = 1, labels = FALSE, tck = -0.03)
abline(h = 0, lty = "dotted")
abline(v = 0, lty = "dotted")
points(NMDS.plot, "sites", cex=0.8, pch=16, col=MDS.colors)
ordiellipse(NMDS.plot, groups=NMDS.df.R14$Period:NMDS.df.R14$dom, draw="polygon", col=MDS.colors, alpha=150, kind="sd", conf=0.8)
legend("topright", legend=MDS.lev.leg, inset=c(-0.05, -0.02), x.intersp=0.8, cex=0.6, pch=16, col=MDS.colors, pt.cex=0.7, bty="n")
par.new=T
plot(vars, p.max=0.05, cex=0.8, lwd=1)

dev.copy(pdf, "figures/NMDS_R14.pdf", height=4, width=5, useDingbats=FALSE)
dev.off() 

```
```{r}

par(mfcol=c(2,1), mar=c(4,4,1,1))


###### plot it
MDS.levels<-levels(NMDS.df.Lil$Period:NMDS.df.Lil$dom)
MDS.lev.leg<-c("C-Bleach '14", "D-Bleach '14", "C-Recov '15", "D-Recov '15", "C-Bleach '15", "D-Bleach '15", "C-Recov '16", "D-Recov '16")
MDS.colors<-c("lightpink", "firebrick2", "slategray1", "dodgerblue2","khaki", "darkorange", "mediumseagreen", "olivedrab")

NMDS.plot<-ordiplot(NMDS.Lil, type="n", main=substitute(paste(italic("Lilipuna"))), cex.main=1, display="sites", cex.lab=0.8, cex.axis=0.8, ylim=c(-0.4, 0.4), xlim=c(-0.6, 0.5))
axis(side = 1, labels = FALSE, tck = -0.03)
abline(h = 0, lty = "dotted")
abline(v = 0, lty = "dotted")
points(NMDS.plot, "sites", cex=0.8, pch=16, col=MDS.colors)
ordiellipse(NMDS.plot, groups=NMDS.df.Lil$Period:NMDS.df.Lil$dom, draw="polygon", col=MDS.colors, alpha=150, kind="sd", conf=0.8)
legend("topright", legend=MDS.lev.leg, inset=c(0.54, -0.02), x.intersp=0.8, y.intersp=0.4, cex=0.6, pch=16, col=MDS.colors, pt.cex=0.7, bty="n")
par.new=T
plot(vars, p.max=0.05, cex=0.8, lwd=1)

###### plot it
MDS.levels<-levels(NMDS.df.R14$Period:NMDS.df.R14$dom)

NMDS.plot<-ordiplot(NMDS.R14, type="n", main=substitute(paste(italic("Reef 14"))), cex.main=1, display="sites", cex.lab=0.8, cex.axis=0.8, ylim=c(-0.4, 0.4), xlim=c(-0.6, 0.5))
axis(side = 1, labels = FALSE, tck = -0.03)
abline(h = 0, lty = "dotted")
abline(v = 0, lty = "dotted")
points(NMDS.plot, "sites", cex=0.8, pch=16, col=MDS.colors)
ordiellipse(NMDS.plot, groups=NMDS.df.R14$Period:NMDS.df.R14$dom, draw="polygon", col=MDS.colors, alpha=150, kind="sd", conf=0.8)
par.new=T
plot(vars, p.max=0.05, cex=0.8, lwd=1)

dev.copy(pdf, "figures/NMDS_LilR14.pdf", height=7, width=5, useDingbats=FALSE)
dev.off() 

```


```{r, eval=FALSE}
### symbiodinium data


###############################
## making figures, tables, analyses

###############################
# make a table by dominant symb

df

str(df)
print(levels(df$Period))

# 4 symbiont categories
symbcomp=table(df$syms, df$Period:df$Site) 
prop.table(symbcomp, margin = 2)

# 2 symbiont categories
domsymb=table(df$dom, df$Period:df$Site)
prop.table(domsymb, margin = 2)


# to specify order in figure, can use... ...(prop.table(symbcomp[,c(1,3,2,4)], 3,4,1,2,5,6
barplot(prop.table(symbcomp, margin = 2), col = c("slategray4", "slategray2", "dodgerblue", "darkturquoise"), main= "2014 - 2016 qPCR", xlab = "Site and Status", cex.names=0.65, ylab = "Proportion of Symbiont Types")
legend("topright", cex=0.8, bty="n", legend=c("C","CD", "D", "DC"), fill=c("slategray4", "slategray2", "dodgerblue", "darkturquoise")) #inset = c(-.25, 0), xpd = NA, x.intersp=0.1, y.intersp=0.7)

dev.copy(pdf, "Figures/symb_4 levels.pdf", encod="MacRoman", height=6, width=14)
dev.off()


#####################
## figures for 2014-2016 dominant symbiont composition figure (2 categories)

# to sepcify order in figure, can use... ...(prop.table(symbcomp[,c(1,3,2,4)],
barplot(prop.table(domsymb, margin = 2), col = c("slategray4", "darkturquoise"), main= "2014 - 2015 qPCR", xlab = "Site and Status", cex.names=0.65, ylab = "Proportion of Symbiont Types")
legend("topright", legend=c("C", "D"), bty="n", fill=c("slategray4", "darkturquoise"), inset = c(-.23, 0), xpd = NA, x.intersp=0.1, y.intersp=0.7)

dev.copy(pdf, "Figures/symb_2 levels.pdf", encod="MacRoman", height=6, width=14)
dev.off()

#Chi Squared test for independence
chisq.test(symbcomp) # p<0.01, accept Ha that symcomb IS related to events
chisq.test(domsymb) # p=0.1, accept Ho that domsymb is NOT related to events

prop.table(symbcomp, margin = 2)
par(mar=c(3, 4, 2, 6))


# to specify order in figure, can use... ...(prop.table(symbcomp[,c(1,3,2,4)],
barplot(prop.table(symbcomp, margin = 2), col = c("slategray4", "slategray2", "dodgerblue", "darkturquoise"), main= "2015 - 2016 qPCR", xlab = "Event and Site", cex.names=0.65, ylab = "Proportion of Symbiont Types")
legend("topright", legend=c("C","CD", "D", "DC"), fill=c("slategray4", "slategray2", "dodgerblue", "darkturquoise"))
#inset = c(-.15, 0), xpd = NA)

##########
prop.table(domsymb, margin = 2)
par(mar=c(3, 4, 2, 6))

barplot(prop.table(domsymb, margin = 2), col = c("slategray4", "darkturquoise"), main= "2015 - 2016 qPCR", xlab = "Site and Status", cex.names=0.65, ylab = "Proportion of Symbiont Types")
legend("topright", legend=c("C", "D"), fill=c("slategray4", "darkturquoise"), inset = c(-.15, 0), xpd = NA)
```
